Online Tuning of EASY-Backfilling using Queue Reordering Policies.

### 2018-01-12 paper/paper.tex
paper fix

Most common work in the literature consider \textit{a posteriori} optimization
of a scheduling algorithm on a (actual) data set. The increasing amount and
diversity of data generated by large scale platforms motivate the community to
study more sophisticated policies that could leverage these data. During the
last few years, there was an explosion in the number of works at the interface
of the HPC and Big Data fields, dealing with learning algorithms. Recent
sophisticated approaches developed \textit{ad hoc} supervised machine learning
algorithms for reducing uncertainty in problem parameters (e.g.,\ job execution
times ~\cite{learningruntimes}, memory requirements, etc.). However, only few studies
evaluate the impact of such techniques on the performances of the resource
manager.

@@

Both of these methods are evaluated with a specifically developed simulator,
which is open-sourced ~\cite{ocst}.

@@

Parallel job scheduling is an old studied theoretical
problem (~\cite{Frachtenberg:2009:JSS:1692356,Feitelson:2004:PJS:2128864.2128865})
whose practical ramifications, varying hypotheses, and inherent uncertainty of
the problem applied to the HPC field have driven practitioners to use simple
heuristics (and researchers to study their behavior). The two most popular
heuristics for HPC platforms are EASY ~\cite{easy} and
Conservative ~\cite{Mu'alem:2001:UPW:380314.380315} backfilling.

While Conservative Backfilling offers many advantages ~\cite{bfchar}, it has a
significant computational overhead and is more complex, which mainly explains
why many of the machines of the top500 ranking ~\cite{top500} rather use a
variant of EASY-Backfilling instead. There is a large body of work seeking to
improve/tune EASY. Indeed, while the basic mechanism is used by some actual
resource and job management software (most notably SLURM ~\cite{SLURMdocSCHED}),
this is seldom done without fine tuning by system administrators.

@@

The dynP scheduler ~\cite{streit_selftuning_2002} suggests a systematic method
to tuning these queues. The algorithm requires simulated scheduling runs at
decision time, therefore the time-to-decision of the scheduling is higher than
that of EASY. The adaptation of EASY was studied under a different framework
than ours in ~\cite{talby2005improving}. Other works study the reservation
mechanism ~\cite{srinivasan2002selective}. Online portfolio approaches are also
studied for the cloud setting ~\cite{deng2013exploring}.

@@

There was a recent focus on leveraging the huge amount of data available in
large scale computing platforms to improve system performance. Some
works use collaborative filtering to co-locate tasks in clouds by estimating
application interference ~\cite{7516031}. Some others are closer to the
application level. For instance,  ~\cite{fmodeling} uses binary classification
to distinguish benign memory faults from application errors so as to execute
recovery algorithms.

Several works use similar techniques in the context of HPC, in
particular ~\cite{Tsafrir_easypp_2005,learningruntimes}, hoping that better job
runtime estimations should improve the scheduling ~\cite{chiang_impact_2002}.
Some algorithms estimate runtime distributions model and choose jobs using
probabilistic integration procedures ~\cite{Nissimov2008}. However, these works
do not address the duality between the cumulative and maximal scheduling costs,
as mentioned in ~\cite{learningruntimes}. While these previous works intend to
estimate uncertain parameters, we consider here a more pragmatic approach,
that directly learns a good scheduling policy from a given policy space.

Existing work ~\cite{jsspp17} takes this approach offline, by
splitting workload data in a training/testing fashion. This work studies
EASY-Backfilling and shows that while the relative performance of some well-known
priority orders for starting jobs differs between workloads, it is relatively
stable over time for a given workload.

@@

A multi-armed bandit (MAB) problem is a sequential allocation problem with
partially observable rewards ~\cite{bubnow}. At every round an action, called an
"arm" in the literature, must be chosen from a fixed set and the corresponding
reward is observed. The goal of a MAB algorithm is to maximize the total reward
obtained in a successive number of rounds. This is essentially achieved using a
combination of explorative actions, which help estimate the quality of each
arm, and of exploitative actions, which leverage the current estimates by using
a good arm. This problem can be addressed using various hypotheses. Most common
are the (original) stochastic setting and the adversarial setting of regret
minimization. See ~\cite{thompson} for the original work on the stochastic
setting and ~\cite{Auer2002} for the ''Upper Confidence Bound`` family of
algorithms. To the best of our knowledge, the work of ~\cite{Banos} is the
earliest for the adversarial setting; see ~\cite{nonstoch} for the
``Exponential-weight'' family of algorithms. We refer to ~\cite{bubnow} for a
comprehensive review of the field. While these algorithms bound the cumulative
difference in loss to the best arm (the regret), they have functional
constraints such as the fact that the rewards should be contained in a range.
The simple heuristic called \textit{Epsilon-Greedy} introduced
in ~\cite{Auer2002} does not have this requirement. For this reason, this
heuristic will be used in this paper and is described in
Subsection~\ref{sub:bandit}.

@@

The selection of the job to run is performed according to a scheduling policy,
which determines the order the jobs are executed. EASY-Backfilling is
the most widely used policy due to its simple and robust implementation and
known benefits such as high system utilization ~\cite{easy}. This strategy has
no worst case guarantee beyond the absence of starvation (i.e.\ every job will
be scheduled at some moment).

@@

\begin{enumerate}
  \item FCFS: First-Come First-Serve, which is the most commonly used policy ~\cite{easy}.
  \item LCFS: Last-Come First-Serve.
  \item SPF: Smallest estimated Processing time $\widetilde{p_{j}}$ First ~\cite{bfchar}.
  \item LPF: Longest estimated Processing time First.
  \item LQF: Largest resource requirement $q_j$ First.
  \item SQF: Smallest resource requirement First.
  \item LEXP: Largest Expansion Factor First ~\cite{bfchar}, where the expansion
    factor is defined as follows:
  \begin{equation} \frac{wait_j + \widetilde{p_j}}{\widetilde{p_j}} \end{equation}
    where $wait_j$ is the time job $j$ has been waiting until the time at which the decision is taken.
    This corresponds the the value of the stretch(or slowdown) where the user runtime estimate
    is used.
  \item SEXP: Smallest Expansion Factor First 
  \item LRF: Largest Ratio $\frac{\widetilde{p_j}}{q_j}$ First
  \item SRF: Smallest Ratio First
  \item LAF: Largest Area $ \widetilde{p_j} \times q_j$ First (often called "Total Resources")
  \item SAF: Smallest Area First
\end{enumerate}

@@

We further assume that there is a certain regularity among periods, i.e.\ that
the distribution of the job submission process does not radically differ
between consecutive periods. This assumption is validated in the study
presented in ~\cite{jsspp17}. It further entails that the behavior of a policy
during the previous periods reflects its behavior on the current one, so that
the selection of a policy can be based on its past behavior. However, there may
be some variability between different periods for certain cost
metrics ~\cite{feitelson2001metrics}. This is illustrated in
Figure~\ref{fig:variability} which displays the average waiting time of various
policies for the KTH-SP2 trace (see~\ref{sub:traces} for a description of the
workloads) using weekly periods encompassing one year. As one can note, the
average waiting time varies a lot from one week to the other, for all 7
policies considered. This indicates that when the cost metric is averaged
over different periods, there is a trade-off to find in between longer periods
that would somehow limit the variability, and shorter ones that yield more
values for the estimation. We will come back to this issue below.

The selection of a policy is reminiscent of reinforcement learning, where
solving a search and inference problem in a (perhaps restrictive) policy space
is easier than in the space of the original problem. It is important to note,
however, that a pure reinforcement learning approach is difficult to develop in
our context. Indeed, while we have outlined a reduced action space, the state
space to consider is infinite. This is not prohibitive \textit{per se}, as
modern methods ~\cite{aprl} can bypass dimensionality issues via function
approximation. However, these methods rely on large amounts of batch data, and
we are facing an online problem with no pre-existing data\footnote{The
reinforcement learning literature uses the term ``\textit{on-policy}'' for this
setting.}. We rely here on simpler, yet we believe more effective, strategies
to solve this problem.  These strategies are applied online and rely on exact
simulation, noisy simulation and $\epsilon$-greedy bandit exploration.

@@

We take another step to evaluate how the simulation strategy for selecting
policies would work in a batch scheduler. Simulation of real systems can be
imprecise. While existing work reports values as low as 2\% imprecision we take
a conservative approach and consider imprecision up to of 20\%. We randomly
introduce multiplicative noise in the estimate of the policy cost defined by
Eq.~\eqref{eq:cost-exact} by rescaling it, either down or up:
%
\begin{equation} \label{eq:cost-noisy} w_{\mbox{noisy}}(t;P_i) = \rho_t^i
\sum_{j=1}^{l(\Delta_t)} \, \mbox{Wait}_j^{P_i} \end{equation}
%
where $\rho_t^i$ is uniformly sampled in the interval $[0.80,1.20]$, thus
adding a +/- 20\% multiplicative noise on the waiting time estimated by the
simulator.  This value is taken as to be sufficiently larger than values
previously reported, see e.g. ~\cite{dutot:hal-01333471} which reports 2
percent variations of the simulated metric compared to the metric on the real
systems.  The overall cost is then defined in the same way as before, leading
to:
%
\begin{equation} \label{eq:tot-cost-noisy} w_{\mbox{noisy}}(\rightarrow T;P_i) =
\sum_{t=0}^{T-1} \lambda^{T-1-t} w_{\mbox{noisy}}(t;P_i) \end{equation}
%
As before the policy that minimizes the above
cost is selected for the period $\Delta_T$.

@@

Thus, we explore here a more efficient strategy that dispenses with estimating
the cost of all policies at a given time and only updates the cost of the
policy that is currently being used. This strategy makes use of the
$\epsilon$-greedy exploration method, standard in reinforcement learning
~\cite{Watkins:1989} and bandit problems ~\cite{Auer2002}, to trade-off
exploitation and exploration, and relies, as before, on past estimates of the
cost to select the best policy in the exploitation mode.

@@

Figure~\ref{fig:bandit} illustrates this process. Several studies study the
decay of $\epsilon$ over time, the exploration being less important once the
estimates for the different policies are reliable ~\cite{Tokic:2010}. This
strategy has however not been beneficial in our case. We believe this is due to
the properties of the traces we use, as explained in
Section~\ref{sec:experiments}. The $\epsilon$-greedy strategy is known to be
outperformed in terms of regret by other classical algorithms such as the UCB
or EXP family of algorithms (see ~\cite{bubnow}). See ~\cite{bubnow} for a precise definition and
thorough treatment of this subject. These algorithms however all rely on
pre-normalized rewards. Here, the scale of $w_{\epsilon}(t;P_i)$ is not known
in advance. While the usual doubling tricks can be devised to go around
sequential issues of scale with good asymptotical performance, they ruin
performance in practice. Accordingly, we only perform experiments using the
simple $\epsilon$-greedy method.

@@

The experimental validation performed in this work uses a mixture of small and
large systems from different periods. These platforms are all targeted by this
work, since they all are homogenous systems that run a batch scheduler.
Table~\ref{tab:logs} presents the 7 logs, which can all be obtained from the
Parallel Workload Archive ~\cite{Feitelson20142967}. We use the
``cleaned'' ~\cite{pwa-clean} version of the logs as per the archive. We impose
an additional filtering step to the workloads\footnote{This filtering step is
available in the reproducible workflow ~\cite{repro} as shell script
\lstinline[basicstyle=\ttfamily\color{blue}]|misc/strong\_filter|} in order to
clean erroneous data:

@@

The choice of a simulator is a critical part of experimental validation that
raises a trade-off between precision and runtime. High precision can be
obtained by carefully modeling the platform and its network topology, or
extracting information from the jobs from their sources or post-mortem logs.
This is the approach used by high-fidelity batch scheduling simulators such as
BatSim ~\cite{batsim}. In the present work, the focus is on studying the
EASY-Backfilling mechanism in itself, without addressing the allocation
problem. Moreover, the experimental protocol used here requires many
simulation runs. Therefore, we set the precision/runtime trade-off at the
point which minimizes simulation runtimes. We discard all topological
information relative to the platform and use the job processing times of the
original workloads. In this setup, the processors are considered to be
undistinguishable from each other, and jobs can be discontinuously mapped to
any available processor on the system. During this work, we developed a
lightweight simulator ~\cite{ocst} and now release it to the community along with
this article. See the reproducibility paragraph below for more information.
With this simulator, we are able to replay EASY-FCFS on the CEA-CURIE trace in
33 seconds with a machine equipped with an Intel(R) Core(TM) i5-4310U CPU @
2.00GHz. As a comparison, the Batsim simulator with network communication
modeling disabled and no resource contiguity takes more than 7 hours to replay
the same trace on the same hardware. The complete set of simulations presented
below is executed on a single Dell PowerEdge R730 in 22 hours, including input
preparation and statistical analysis code.

@@

Comparing algorithms in batch scheduling is made especially difficult by the
fluctuations in commonly used metrics ~\cite{feitelson1998metrics}. The variability in
performance is known to outrank the available sample sizes: \textit{It is not
enough to simply replay an algorithm using a workload trace}. Indeed, as
presented in Figure~\ref{fig:variability}, the objective function we are using
in this work is subject to high variability. In our experience, randomly
shuffling months in a trace at random is enough to generate contradictory
conclusions about which algorithm is better. This is a known
problem ~\cite{flurries} that motivated decades of research in workload
modeling ~\cite{feitbook}. Workload models have drawbacks in that they usually
lose the details of the workloads. This is problematic due to the non-linear
aspect of the job scheduling process. This is discussed
in ~\cite{feitresampling}, an approach that attempts to solve this problem by
resampling data directly from the original traces. In this work, we use a
simple implementation of these ideas. More precisely, we want to sample the job
submission process, which is achieved in the following manner: (a) For every
week in a resampled trace, we iterate over every existing user in the original
trace; (b) For each of these users, we choose a week of the original trace
uniformly at random; (c) The jobs from this user in this original week are
chosen to be present in the week being resampled. This simple mechanism enables
us to preserve the dependency within a week. Although the system state may not
be independent from one week to an other, as illustrated in~\cite{jsspp17},
it is reasonable
to assume that the system is in a stationary state.~\cite{jsspp17} and assume for these experiments that the system is in a
stationary state. We do provide guidelines as how to incorporate tracking
into our algorithms in Section~\ref{sec:online} by introducing a decaying
discount factor $\lambda$. In our experiments, we use stationary resampled data, 
and therefore no decay is necessary: $\lambda$ here is set to $1$.

@@

For each of the 7 workloads
considered, we use the original data to resample 60 traces, each with a length
of two years. For each resampled trace, we run simulations using the
EASY($P,\Theta$) scheduling heuristic. We fix the value of the job waiting time
threshold $\Theta$ at 40 hours. The choice of this default
stems from the analysis developed in ~\cite{jsspp17}. Essentially, this value is high
enough for queue reorderings to be leveraged while being low enough to provide
the functional requirement of preventing starvation. We simulate
EASY($P,\Theta=40h$) using the following strategies:

@@

We publish a declarative workflow ~\cite{repro} based on our
simulator ~\cite{ocst} that runs experiments and generates all the figures from
this paper.

There are various approaches for ensuring computational experiments are replicable,
among which distributing complete operating system images, containers, or
packaging software. As our experiments are CPU bound, we decide to opt for the
software packaging approach ~\cite{nix}. This allows to replicate the experiments on
clusters where virtualization is not available or kernels are too old for
containers and decreases simulation runtime.

@@

All the code associated with this work is released under the ISC ~\cite{isc}
license.
Running the experiments can be done
on any platform equipped with the Nix package manager by running at the root of
the extracted archive:

@@

All data from this paper were obtained from the Swf Parallel Workloads
Archive ~\cite{Feitelson20142967}.

 All experiments are tied together using zymake ~\cite{zymake}, a minimalistic
 workflow system designed for computational experiments\footnote{The zymake system
 is also packaged by our nix expressions.}. This system is
 analogous to a traditional build system with added workflow capabilities. The
 entire workflow that generates this article from the input data is contained
 in a single \lstinline[basicstyle=\ttfamily\color{blue}]|zymakefile| to be
 found at the root of the main archive. This workflow is composed of the
 following steps:

@@

   \item Simulation. This step runs the lightweight ocaml backfilling
     simulator specially written for this work. This simulator is made
     available under the ISC ~\cite{isc} license as a persistent zenodo
     archive at ~\cite{ocst}.

>>>>

Most common work in the literature consider \textit{a posteriori} optimization
of a scheduling algorithm on a (actual) data set. The increasing amount and
diversity of data generated by large scale platforms motivate the community to
study more sophisticated policies that could leverage these data. During the
last few years, there was an explosion in the number of works at the interface
of the HPC and Big Data fields, dealing with learning algorithms. Recent
sophisticated approaches developed \textit{ad hoc} supervised machine learning
algorithms for reducing uncertainty in problem parameters (e.g.,\ job execution
times~\cite{learningruntimes}, memory requirements, etc.). However, only few studies
evaluate the impact of such techniques on the performances of the resource
manager.

@@

Both of these methods are evaluated with a specifically developed simulator,
which is open-sourced~\cite{ocst}.

@@

Parallel job scheduling is an old studied theoretical
problem (~\cite{Frachtenberg:2009:JSS:1692356,Feitelson:2004:PJS:2128864.2128865})
whose practical ramifications, varying hypotheses, and inherent uncertainty of
the problem applied to the HPC field have driven practitioners to use simple
heuristics (and researchers to study their behavior). The two most popular
heuristics for HPC platforms are EASY~\cite{easy} and
Conservative~\cite{mualem_utilization_2001} backfilling.

While Conservative Backfilling offers many advantages~\cite{bfchar}, it has a
significant computational overhead and is more complex, which mainly explains
why many of the machines of the top500 ranking~\cite{top500} rather use a
variant of EASY-Backfilling instead. There is a large body of work seeking to
improve/tune EASY. Indeed, while the basic mechanism is used by some actual
resource and job management software (most notably SLURM~\cite{SLURMdocSCHED}),
this is seldom done without fine tuning by system administrators.

@@

The dynP scheduler~\cite{streit_selftuning_2002} suggests a systematic method
to tuning these queues. The algorithm requires simulated scheduling runs at
decision time, therefore the time-to-decision of the scheduling is higher than
that of EASY. The adaptation of EASY was studied under a different framework
than ours in~\cite{talby2005improving}. Other works study the reservation
mechanism~\cite{srinivasan2002selective}. Online portfolio approaches are also
studied for the cloud setting~\cite{deng2013exploring}.

@@

There was a recent focus on leveraging the huge amount of data available in
large scale computing platforms to improve system performance. Some
works use collaborative filtering to co-locate tasks in clouds by estimating
application interference~\cite{7516031}. Some others are closer to the
application level. For instance, ~\cite{fmodeling} uses binary classification
to distinguish benign memory faults from application errors so as to execute
recovery algorithms.

Several works use similar techniques in the context of HPC, in
particular~\cite{Tsafrir_easypp_2005,learningruntimes}, hoping that better job
runtime estimations should improve the scheduling~\cite{chiang_impact_2002}.
Some algorithms estimate runtime distributions model and choose jobs using
probabilistic integration procedures~\cite{Nissimov2008}. However, these works
do not address the duality between the cumulative and maximal scheduling costs,
as mentioned in~\cite{learningruntimes}. While these previous works intend to
estimate uncertain parameters, we consider here a more pragmatic approach,
that directly learns a good scheduling policy from a given policy space.

Existing work~\cite{jsspp17} takes this approach offline, by
splitting workload data in a training/testing fashion. This work studies
EASY-Backfilling and shows that while the relative performance of some well-known
priority orders for starting jobs differs between workloads, it is relatively
stable over time for a given workload.

@@

A multi-armed bandit (MAB) problem is a sequential allocation problem with
partially observable rewards~\cite{bubnow}. At every round an action, called an
"arm" in the literature, must be chosen from a fixed set and the corresponding
reward is observed. The goal of a MAB algorithm is to maximize the total reward
obtained in a successive number of rounds. This is essentially achieved using a
combination of explorative actions, which help estimate the quality of each
arm, and of exploitative actions, which leverage the current estimates by using
a good arm. This problem can be addressed using various hypotheses. Most common
are the (original) stochastic setting and the adversarial setting of regret
minimization. See~\cite{thompson} for the original work on the stochastic
setting and~\cite{Auer2002} for the ''Upper Confidence Bound`` family of
algorithms. To the best of our knowledge, the work of~\cite{Banos} is the
earliest for the adversarial setting; see~\cite{nonstoch} for the
``Exponential-weight'' family of algorithms. We refer to~\cite{bubnow} for a
comprehensive review of the field. While these algorithms bound the cumulative
difference in loss to the best arm (the regret), they have functional
constraints such as the fact that the rewards should be contained in a range.
The simple heuristic called \textit{Epsilon-Greedy} introduced
in~\cite{Auer2002} does not have this requirement. For this reason, this
heuristic will be used in this paper and is described in
Subsection~\ref{sub:bandit}.

@@

The selection of the job to run is performed according to a scheduling policy,
which determines the order the jobs are executed. EASY-Backfilling is
the most widely used policy due to its simple and robust implementation and
known benefits such as high system utilization~\cite{easy}. This strategy has
no worst case guarantee beyond the absence of starvation (i.e.\ every job will
be scheduled at some moment).

@@

\begin{enumerate}
  \item FCFS: First-Come First-Serve, which is the most commonly used policy~\cite{easy}.
  \item LCFS: Last-Come First-Serve.
  \item SPF: Smallest estimated Processing time $\widetilde{p_{j}}$ First~\cite{bfchar}.
  \item LPF: Longest estimated Processing time First.
  \item LQF: Largest resource requirement $q_j$ First.
  \item SQF: Smallest resource requirement First.
  \item LEXP: Largest Expansion Factor First~\cite{bfchar}, where the expansion
    factor is defined as follows:
  \begin{equation} \frac{wait_j + \widetilde{p_j}}{\widetilde{p_j}} \end{equation}
    where $wait_j$ is the time job $j$ has been waiting until the time at which the decision is taken.
    This corresponds the the value of the stretch(or slowdown) where the user runtime estimate
    is used.
  \item SEXP: Smallest Expansion Factor First 
  \item LRF: Largest Ratio $\frac{\widetilde{p_j}}{q_j}$ First
  \item SRF: Smallest Ratio First
  \item LAF: Largest Area $ \widetilde{p_j} \times q_j$ First (often called "Total Resources")
  \item SAF: Smallest Area First
\end{enumerate}

@@

We further assume that there is a certain regularity among periods, i.e.\ that
the distribution of the job submission process does not radically differ
between consecutive periods. This assumption is validated in the study
presented in~\cite{jsspp17}. It further entails that the behavior of a policy
during the previous periods reflects its behavior on the current one, so that
the selection of a policy can be based on its past behavior. However, there may
be some variability between different periods for certain cost
metrics~\cite{feitelson2001metrics}. This is illustrated in
Figure~\ref{fig:variability} which displays the average waiting time of various
policies for the KTH-SP2 trace (see~\ref{sub:traces} for a description of the
workloads) using weekly periods encompassing one year. As one can note, the
average waiting time varies a lot from one week to the other, for all 7
policies considered. This indicates that when the cost metric is averaged
over different periods, there is a trade-off to find in between longer periods
that would somehow limit the variability, and shorter ones that yield more
values for the estimation. We will come back to this issue below.

The selection of a policy is reminiscent of reinforcement learning, where
solving a search and inference problem in a (perhaps restrictive) policy space
is easier than in the space of the original problem. It is important to note,
however, that a pure reinforcement learning approach is difficult to develop in
our context. Indeed, while we have outlined a reduced action space, the state
space to consider is infinite. This is not prohibitive \textit{per se}, as
modern methods~\cite{aprl} can bypass dimensionality issues via function
approximation. However, these methods rely on large amounts of batch data, and
we are facing an online problem with no pre-existing data\footnote{The
reinforcement learning literature uses the term ``\textit{on-policy}'' for this
setting.}. We rely here on simpler, yet we believe more effective, strategies
to solve this problem.  These strategies are applied online and rely on exact
simulation, noisy simulation and $\epsilon$-greedy bandit exploration.

@@

We take another step to evaluate how the simulation strategy for selecting
policies would work in a batch scheduler. Simulation of real systems can be
imprecise. While existing work reports values as low as 2\% imprecision we take
a conservative approach and consider imprecision up to of 20\%. We randomly
introduce multiplicative noise in the estimate of the policy cost defined by
Eq.~\eqref{eq:cost-exact} by rescaling it, either down or up:
%
\begin{equation} \label{eq:cost-noisy} w_{\mbox{noisy}}(t;P_i) = \rho_t^i
\sum_{j=1}^{l(\Delta_t)} \, \mbox{Wait}_j^{P_i} \end{equation}
%
where $\rho_t^i$ is uniformly sampled in the interval $[0.80,1.20]$, thus
adding a +/- 20\% multiplicative noise on the waiting time estimated by the
simulator.  This value is taken as to be sufficiently larger than values
previously reported, see e.g.~\cite{dutot:hal-01333471} which reports 2
percent variations of the simulated metric compared to the metric on the real
systems.  The overall cost is then defined in the same way as before, leading
to:
%
\begin{equation} \label{eq:tot-cost-noisy} w_{\mbox{noisy}}(\rightarrow T;P_i) =
\sum_{t=0}^{T-1} \lambda^{T-1-t} w_{\mbox{noisy}}(t;P_i) \end{equation}
%
As before the policy that minimizes the above
cost is selected for the period $\Delta_T$.

@@

Thus, we explore here a more efficient strategy that dispenses with estimating
the cost of all policies at a given time and only updates the cost of the
policy that is currently being used. This strategy makes use of the
$\epsilon$-greedy exploration method, standard in reinforcement learning
~\cite{Watkins:1989} and bandit problems~\cite{Auer2002}, to trade-off
exploitation and exploration, and relies, as before, on past estimates of the
cost to select the best policy in the exploitation mode.

@@

Figure~\ref{fig:bandit} illustrates this process. Several studies study the
decay of $\epsilon$ over time, the exploration being less important once the
estimates for the different policies are reliable~\cite{Tokic:2010}. This
strategy has however not been beneficial in our case. We believe this is due to
the properties of the traces we use, as explained in
Section~\ref{sec:experiments}. The $\epsilon$-greedy strategy is known to be
outperformed in terms of regret by other classical algorithms such as the UCB
or EXP family of algorithms (see~\cite{bubnow}). See~\cite{bubnow} for a precise definition and
thorough treatment of this subject. These algorithms however all rely on
pre-normalized rewards. Here, the scale of $w_{\epsilon}(t;P_i)$ is not known
in advance. While the usual doubling tricks can be devised to go around
sequential issues of scale with good asymptotical performance, they ruin
performance in practice. Accordingly, we only perform experiments using the
simple $\epsilon$-greedy method.

@@

The experimental validation performed in this work uses a mixture of small and
large systems from different periods. These platforms are all targeted by this
work, since they all are homogenous systems that run a batch scheduler.
Table~\ref{tab:logs} presents the 7 logs, which can all be obtained from the
Parallel Workload Archive~\cite{Feitelson20142967}. We use the
``cleaned''~\cite{pwa-clean} version of the logs as per the archive. We impose
an additional filtering step to the workloads\footnote{This filtering step is
available in the reproducible workflow~\cite{repro} as shell script
\lstinline[basicstyle=\ttfamily\color{blue}]|misc/strong\_filter|} in order to
clean erroneous data:

@@

The choice of a simulator is a critical part of experimental validation that
raises a trade-off between precision and runtime. High precision can be
obtained by carefully modeling the platform and its network topology, or
extracting information from the jobs from their sources or post-mortem logs.
This is the approach used by high-fidelity batch scheduling simulators such as
BatSim~\cite{batsim}. In the present work, the focus is on studying the
EASY-Backfilling mechanism in itself, without addressing the allocation
problem. Moreover, the experimental protocol used here requires many
simulation runs. Therefore, we set the precision/runtime trade-off at the
point which minimizes simulation runtimes. We discard all topological
information relative to the platform and use the job processing times of the
original workloads. In this setup, the processors are considered to be
undistinguishable from each other, and jobs can be discontinuously mapped to
any available processor on the system. During this work, we developed a
lightweight simulator~\cite{ocst} and now release it to the community along with
this article. See the reproducibility paragraph below for more information.
With this simulator, we are able to replay EASY-FCFS on the CEA-CURIE trace in
33 seconds with a machine equipped with an Intel(R) Core(TM) i5-4310U CPU @
2.00GHz. As a comparison, the Batsim simulator with network communication
modeling disabled and no resource contiguity takes more than 7 hours to replay
the same trace on the same hardware. The complete set of simulations presented
below is executed on a single Dell PowerEdge R730 in 22 hours, including input
preparation and statistical analysis code.

@@

Comparing algorithms in batch scheduling is made especially difficult by the
fluctuations in commonly used metrics~\cite{feitelson1998metrics}. The variability in
performance is known to outrank the available sample sizes: \textit{It is not
enough to simply replay an algorithm using a workload trace}. Indeed, as
presented in Figure~\ref{fig:variability}, the objective function we are using
in this work is subject to high variability. In our experience, randomly
shuffling months in a trace at random is enough to generate contradictory
conclusions about which algorithm is better. This is a known
problem~\cite{flurries} that motivated decades of research in workload
modeling~\cite{feitbook}. Workload models have drawbacks in that they usually
lose the details of the workloads. This is problematic due to the non-linear
aspect of the job scheduling process. This is discussed
in~\cite{feitresampling}, an approach that attempts to solve this problem by
resampling data directly from the original traces. In this work, we use a
simple implementation of these ideas. More precisely, we want to sample the job
submission process, which is achieved in the following manner: (a) For every
week in a resampled trace, we iterate over every existing user in the original
trace; (b) For each of these users, we choose a week of the original trace
uniformly at random; (c) The jobs from this user in this original week are
chosen to be present in the week being resampled. This simple mechanism enables
us to preserve the dependency within a week. Although the system state may not
be independent from one week to an other, as illustrated in~\cite{jsspp17},
it is reasonable
to assume that the system is in a stationary state.~\cite{jsspp17} and assume for these experiments that the system is in a
stationary state. We do provide guidelines as how to incorporate tracking
into our algorithms in Section~\ref{sec:online} by introducing a decaying
discount factor $\lambda$. In our experiments, we use stationary resampled data, 
and therefore no decay is necessary: $\lambda$ here is set to $1$.

@@

For each of the 7 workloads
considered, we use the original data to resample 60 traces, each with a length
of two years. For each resampled trace, we run simulations using the
EASY($P,\Theta$) scheduling heuristic. We fix the value of the job waiting time
threshold $\Theta$ at 40 hours. The choice of this default
stems from the analysis developed in~\cite{jsspp17}. Essentially, this value is high
enough for queue reorderings to be leveraged while being low enough to provide
the functional requirement of preventing starvation. We simulate
EASY($P,\Theta=40h$) using the following strategies:

@@

We publish a declarative workflow~\cite{repro} based on our
simulator~\cite{ocst} that runs experiments and generates all the figures from
this paper.

There are various approaches for ensuring computational experiments are replicable,
among which distributing complete operating system images, containers, or
packaging software. As our experiments are CPU bound, we decide to opt for the
software packaging approach~\cite{nix}. This allows to replicate the experiments on
clusters where virtualization is not available or kernels are too old for
containers and decreases simulation runtime.

@@

All the code associated with this work is released under the ISC~\cite{isc}
license.
Running the experiments can be done
on any platform equipped with the Nix package manager by running at the root of
the extracted archive:

@@

All data from this paper were obtained from the Swf Parallel Workloads
Archive~\cite{Feitelson20142967}.

 All experiments are tied together using zymake~\cite{zymake}, a minimalistic
 workflow system designed for computational experiments\footnote{The zymake system
 is also packaged by our nix expressions.}. This system is
 analogous to a traditional build system with added workflow capabilities. The
 entire workflow that generates this article from the input data is contained
 in a single \lstinline[basicstyle=\ttfamily\color{blue}]|zymakefile| to be
 found at the root of the main archive. This workflow is composed of the
 following steps:

@@

   \item Simulation. This step runs the lightweight ocaml backfilling
     simulator specially written for this work. This simulator is made
     available under the ISC~\cite{isc} license as a persistent zenodo
     archive at~\cite{ocst}.
